### 2014-01-20 paper/reply.tex
new reply

\documentclass[12pts]{article}
\begin{document}
Dear Editor, \\

We wish to thank the referee for a detailed report on our paper. \\

The referee expressed 5 major points of concern. Some of them were due
to unclear written sections, other made reference to issues with the
choice of observational data used to compare against the mocks.\\

In this new version we respond to all these concerns. We have
clarified the text where it was needed and performed new tests on the
data. We have also corrected a small error in the presentation of the
results for the angular correlation function. All these modifications
and improvements strengthen the conclusions already
presented. Therefore we resubmit this draft to be considered for
publication in the MNRAS.\\

We wish to emphasize that our paper presents a new method in the context
of the study of LAE at high redshift. Correspondingly, the results are
novel with respect to the already published literature.\\

In what follows we reply to each one of the major points raised by the
referee. The comments by the referee are in boldface. Our reply
follows to each one of the comments.\\

Best regards,\\

Jaime E. Forero-Romero and Julian Mejia-Restrepo\\

{\large \bf Major points}\\

\begin{enumerate}
\item {\bf As the authors are aware, each observational survey for LAEs has
  different selection effects, and some use very different thresholds
  in equivalent width.  Hence you need to use the number density and
  autocorrelation function (ACF) of LAEs from the *same* survey for an
  investigation like this.  You could try to get Yamada's LAE
  positions to calculate the ACF, since you appear to prefer that
  survey for number density estimation.  Or you could pursue joint
  analysis of number density and ACF for smaller surveys e.g.,
  Ouchi's. }\\

  We follow the referee's suggestion and perform the ACF analysis
  using Ouchi's data.

  We made this choice because neither the galaxy catalog
  nor the ACF from Yamada et al. 2012 are available. We contacted
  Dr. Yamada asking for information about the ACF, but we didn't get
  any reply.

\item {\bf 
  The major result claimed that all models have a narrow mass range
  of halos hosting LAEs is guaranteed by the choice of a minimum
  value of $f_{\rm occ}$ of 0.1.  This should be obvious to the authors, as
  there is a degeneracy between $f_{\rm occ}$ and the mass range of halos
  that you have hidden by forcing $f_{\rm occ}$ to values larger than those
  previously claimed in the literature. }

  As the referee highlights, we have found a new set of acceptable
  parameters for the escape fraction $f_{\rm occ}>0.1$. We stress that
  an occupation fraction of $0.1$ is on the upper end of the accepted
  values by Gawiser et al. 2007. At the same time all the mass ranges
  we find acceptable are within the wide mass range already published
  by observers. All this seems to be good news rather a caveat.

  The reason for a narrow mass range $\Delta M$ does not
  seem to be a consequence from the constraint on the occupation
  fraction to be larger than $0.1$. We do not find a clear evidence
  for a degeneracy between the occupation fraction $f_{\rm occ}$ and
  the mass range $\Delta M$. However we would appreciaty a pointer to
  any reference presenting evidence for the opossite.

  However, there is a clear degeneracy between the minimum halo mass
  $M_{\rm min}$ and the occupation fraction.This trend is clearly
  seen on the right panel of Figure 3,4 and 6. This implies that
  exploring very low escape fractions values of the range $0.01$ would
  require building mock catalogs with a minimum mass around the range
  $10^{9.5}$Msun or below which is precisely the minimum halo mass of
  that is resolved in the simulation. This intrinsic limitation on the
  range of values that can be explored is now explicit in the paper.

  All in all, we consider this result for a narrow
  mass range $\Delta M<1.0$ (in many cases $\Delta M<0.5$) a novel
  one, since it has not been mentioned in the literature before.

  \item {\bf 
    The authors cite similar work by Walker Soler et al. 2012 but fail
    to mention in the introduction how the current paper has the same
    goals as that earlier work.  If applied more carefully, their
    approach should be able to improve on the earlier work, but it is
    important to be clear about how similar the papers are and how the
    current one hopes to improve on the previous one. 
  }

The work of Walker-Soler 2012 actually does not pursue a similar goal
as our paper. They take a fixed set of results of a semi-analytical
model to infer the mass of the \emph{descendants} from LAEs at
$z=3.1$. We have removed the reference to that work that might show
that we are improving work they have been done, because that's not the case.

\item {\bf 
Comparing the most heavily clustered of 15 mocks against SSA22 is
  unfair. SSA22 is a ~5 sigma overdensity, which is much rarer than
  1/15. 
  }

We agree with the referee that a comparison of a $\sim 5 $ overdensity
with an average field would not be fair. Therefore we have removed
the comparison from the ACF in the SSA22 field in the new version of
the paper. We only include the ACF comparison against the results of
an average density field from Ouchi 2008.

Related to this point we have added the left panel in Figure 1 and
subsection 2.2.1 to discuss the adequacy of including SSA22 as an
observational benchmark in the number density tests.

In that subsection we emphasize that it is only one of the subfields
in SSA22 that represents at 3-$\sigma$ overdensity. The whole field
has an average density compared to other blank fields.

\item {\bf 
  There seems to be significant confusion between halo occupation
   fraction and Lyman alpha escape fraction.  Only in an incredibly
   simplified model where 100\% of Ly alpha photons escape from LAEs
   and 0\% escape from non-LAEs would the comparison in section 4.1
   make sense.  In a more realistic model, it is not clear how this
   investigation could hope to comment on the escape fraction of Ly
   alpha photons. 
}\\

We agree with the referee in this point.

We have corrected all the mistakes where we meant occupation
fraction but wrote escape fraction. We have also removed the discussion
comparing our results to escape fraction estimates, in spite of the
effort of other authors to link these two different quantities.

\end{enumerate}

Additional suggestions:
\begin{itemize}
\item{\bf Only allowing one LAE per dark matter halo is a simplistic model.  You
could try more complicated HOD models where $f_occ(M)$ is not constant
to see if they make a significant difference in the results.  But
arguing against this by citing the Jose et al. 2013 paper seems
reasonable.  You could also note the lack of close-proximity LAE pairs
noted in papers led by Nick Bond as further evidence against multiple
LAEs occurring in the same halo. }

We would like to add a reference to the work of Nick Bond, but failed
to fine what exact paper makes reference to the lack of
close-proximity LAE pairs.

\item{\bf You could explore a wide range of $f_occ$ down to properly tiny values
such as 0.001 without increasing the run time by indexing it
logarithmically.}

As we mentioned before in the reply to Major Point 2, tiny values of
0.001 can only produce reasonable models for very low mass halos. This
is due to the constraint on the number density distribution. An
occupation fraction of 0.001 requires models with a minimum mass on
the order of $M_{\rm min}=10^{9}$ solar masses, which is well below
the mass resolution of the simulation we use. This comment has been
added to the paper.

\item {\bf Be careful when turning filters into cosmological volumes -
  you should not simply use the FWHM.  See Gronwall et al. 2007 for a detailed
discussion of this.}

We have included a paragraph about this point in Section 4.4.

\item {\bf Be careful when comparing results for halo occupation fraction for
LAEs from a survey with a 190A observed EW threshold to the more
typical 20 A EW rest-frame, which equals 80A observed at z=3. }

We have included a paragraph about this point in Section 4.4

\item {\bf SSA22 is known to be an extraordinary overdensity, as you note.  The
overdense region, and possibly the entire field, will bias the
measured number density of z=3.1 LAEs upwards.  So one should be very
cautious including this field in a study like this one. }

We have included subsection 2.2 to comment on this point.

\item {\bf When using a constant value for $f_{\rm occ}$, you could use just
  $M_{\rm min}$, $M_{\rm max}$ as parameters, choose the value of
  $f_{\rm occ}$ that fits the observed number density (propagating
  uncertainties) and then test the resulting ACF versus data.  This
  would be accurate because a random sub-sample of halos has the same
  clustering as the full sample. }

  We agree that this is a feasible approach. The emphasis we meant in
  our approach was the effect of cosmic variance on the size of
  observed fields. This is why we include the tests on all sub-samples
  and not just one.

\end{itemize}

\end{document}